### 2020-12-11 awr_body.tex
AWR body submission

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}

% Academic Writing
% https://jazyky.fel.cvut.cz/vyuka/RPP/BE9M04AKP/

% How to export to DOCX: Export to PDF and then load the PDF in Word.

\usepackage[disable]{todonotes}

\input{preamble}

\begin{document}

\maketitle

\section{Example body paragraph}

Simplification ordering on terms
influences the proof search in \vampire{} on two levels.
First, the inferences on each clause are limited
to the selected literals.
In each clause,
either a negative literal or all the maximal literals are selected.
The maximality is evaluated
according to the simplification ordering.
% Note: The selection function Total does not use the simplification ordering.
Second, simplification ordering orients some of the equalities
to prevent superposition and equality factoring
from inferring redundant complex conclusions.
In each of these two roles,
the simplification ordering may impact the direction and,
in effect, the length of the proof search.

\end{document}