### 2021-02-16 abstract.tex
Remove an old disabled version of abstract

% !TEX root = main.tex

% AWR: Guidelines (Unit 7, page 5):
% At most 300 words
% Content:
% Purpose of the study
% Research problems
% Basic design of the study
% Summary of interpretations and conclusions

% AWR: Parts of an abstract (checklist) (Unit 7, page 6):
% Motivation
% Problem statement
% Approach
% Results
% Conclusions

State-of-the-art superposition-based theorem provers for \acrlong{fol}
rely on \glspl{sot} to constrain the applicability of inference rules,
which in turn shapes the ensuing search space.
The popular Knuth-Bendix simplification ordering is parameterized by 
\emph{symbol precedence}---a permutation of the predicate and function symbols
of the input problem's signature.
Thus, the choice of precedence has an indirect yet often substantial impact
on the amount of work required to complete a proof search.

\todo{MS: idea maybe to work out more in the introduction: 
We could stress how the impact or the choice of a precedence is \emph{indirect}, 
(as it's already obvious from the explanation here)
all the more interesting one can learn from observing just this indirect impact 
which precendences are good and which are bad!}

This paper describes and evaluates a symbol precedence recommender,
a machine learning system that estimates the best possible precedence
based on observations of prover performance on a set of problems and random precedences.
Using the graph convolutional neural network technology,
the system does not presuppose the problems to be related or share a common signature. 
When coupled with the theorem prover Vampire and evaluated on the \acrshort{tptp} problem library,
the recommender is found to outperform a state-of-the-art heuristic by more than \SI{2}{\percent}.

\keywords{saturation-based theorem proving \and
simplification ordering \and
symbol precedence \and
\acrlong{ml} \and
\acrlong{gcn}}

% Motivation and purpose of the study are excluded because the audience at CADE need not be reminded of the motivation for making the provers faster.