### 2021-02-22 appendix.tex
Move the appendix content into a dedicated file

\section{Graph structure}

\begin{itemize}
\item The graph representation of problem $P$ contains exactly one root node of type \ntype{problem}.
\item Each clause is represented by a \ntype{clause} node connected to the root \ntype{problem} node.
\item Each atom is represented by an \ntype{atom} node (in case the atom is not an equality)
or an \ntype{equality} node (in case the atom is an equality).
For each literal occurrence there is an edge connecting the respective atom to the respective clause.
The type of the edge corresponds to the polarity of the literal:
\epos{} for positive literal and \eneg{} for negative literal.
\item Each \ntype{equality} node is connected to two nodes that represent the operands,
each of which is of type \ntype{term} or \ntype{variable}.
The commutativity of the equality operator is reflected by the fact that the operand edges are not ordered.
\item Each \ntype{atom} node is connected to one \ntype{predicate} node that represents the predicate symbol being applied by this atom.
Note that these edges connect the applications of a predicate symbol across the whole problem.
\item Each \ntype{atom} node is connected to zero or more \ntype{argument} nodes that represent the argument positions of the atom.
\item Each pair of \ntype{argument} nodes that correspond to consecutive argument positions is connected by an edge.
\item Each \ntype{argument} node is connected to a node that represents the argument term,
which is either a \ntype{term} node or a \ntype{variable} node.
\item Each \ntype{term} node is connected to one \ntype{function} node that represents the function symbol being applied by this term.
\item For each variable, there is an edge connecting the \ntype{clause} node of the clause that binds the variable to the \ntype{variable} node that represents the variable.
\end{itemize}

\section{Loss derivative}

The loss is differentiable with respect to the symbol costs:
\begin{align*}
\frac{\partial \loss}{\partial c_i}
&= -\sigmoid(-C(\Better{\PrecBetter}{\PrecWorse}{P})) \cdot k(n) \cdot (\inv{\PrecWorse}_i - \inv{\PrecBetter}_i) \\
&= (p(\Better{\PrecBetter}{\PrecWorse}{P}) - 1) \cdot k(n) \cdot (\inv{\PrecWorse}_i - \inv{\PrecBetter}_i)
\end{align*}

This means that it is possible to backpropagate the loss gradient into the symbol cost model. \todo{MS: tohle bude tezky,
ale ackoliv ML people by tohle uz asi chapali, ATP crowd spis bude
potrebovat obsirnejsi vysvetleni.}

\section{Experiment details}

\begin{figure}[h]
\caption{The dependence of preprocessing time on the number of nodes in the problem graph on 1000 random validation problems}
\label{fig:preprocessing}
\centering
% \includesvg[width=\textwidth]{preprocessing}
% Source: https://docs.google.com/spreadsheets/d/1GujYNEtETpC3jk4iyENLjptv8mDmI5f1ZEhbmwtqnPM/edit#gid=394425425&fvid=659935521
% Experiment: VML-715
\end{figure}