### 2021-03-30 rebuttal.txt
Add a draft of the rebuttal
Reviews can be viewed here:
https://easychair.org/conferences/submission?a=25958734;submission=5327793

We are grateful for the attention and care the reviewers took when considering our paper.

Response to Review 1
--------------------

- An obvious question is how is this work related to the work and methods used for Enigma?

[FB: I'll try to read the Enigma papers and compare our architecture etc. to theirs.]

- Despite the clear improvement over baseline, the percentage gained does not appear
to be too big. It is possible, that the potential effect of better orderings is not
too high, regardless of the optimizations. Can you give some insights for this
line of thought?

The impact of symbol precedence on the proof search is limited,
especially in case it is only used as a tie-breaker in the regular KBO.
[FB: Explain in more detail.]

Response to Review 2
--------------------

- How does your tool behave in the presence of theories (assuming AVATAR is used)?

Both in the training and in the evaluation, AVATAR was disabled.
To avoid confusion, we intend to extend the final version of the paper with an appendix including a detailed description of the Vampire configuration.

- Now Vampire can also handle higher-order problems. Can your technique also improve Vampire's performance in this context?

[FB: Martin, are the precedences compatible with the HOL reasoning in Vampire?]

- How long did it take to do all the training?

In total, the generation of the training data and the training of the model took approximately 8 days.
More precisely, it took approximately 2 and 4 days to generate the training data for the predicate and function precedence recommender, respectively,
and it took approximately 45 and 20 hours to train the predicate and function precedence recommender, respectively.

- What kind of new problems were solved in terms of the rating of the problems?

A total of 234 new problems were solved by the trained recommender when compared to the baseline heuristic.
Among these 234 problems, the maximum rating is 0.88,
the mean rating is approximately 0.36 and the median rating is 0.36.

- How independent are the training and test sets, given that the problems in TPTP are grouped by families that may include extremely similar problems?

In this experiment, we have not attempted to control nor analyze the similarity between the problems in the training and test sets.
Splitting the problem set by the domain or the source of the problems is, however, an interesting idea.

Response to Review 3
--------------------

- Why do you use the number or iterations of the main loop?

The idea behind using the number of iterations of the main loop is that it ensures that the orientation of each pair of precedences is unique –
once we observe that precedence pi is better than precedence rho, we are certain that we will never observe the opposite orientation.
While this approach reduces the amount of noise in the training data,
it is not necessary for the training to succeed.
The point raised by the reviewer is valid and suggests that using wall clock time as the decisive metric, if implemented well, may increase the performance of the trained system.

- You don't describe how many "incomplete" precedence pairs you
have, e.g. pair where one precedence leads to a proof, the other
to a timeout.

[FB: I'll try to extract this from the data. Most likely this information is not available.]

- If I understand you correctly, the graph representation of a problem is fixed. So why don't you filter before generating training examples?

We agree that filtering the problems first by the graph representation size would make the procedure clearer.
The procedure described in the paper was used mainly because it allowed cheaper tuning of the hyperparameters that control the conversion to graphs --
in most of the experiments, we only need to convert a small part of the training and test problems,
and determining whether a graph representation of a problem has at most 100 000 nodes requires performing the conversion to graph.
Filtering by the graph size before splitting into training and test set, once the hyperparameters are tuned, seems to be a reasonable improvement.