### 2021-03-31 rebuttal.txt
minimalni upravky

We are very grateful for the attention and care the reviewers took when considering our paper.

----------- Response to Review 1 -----------

- An obvious question is how is this work related to the work and methods used for Enigma?

ENIGMA [9, Jakubův 2020] differs from our neural precedence recommender especially in the task it deals with:
ENIGMA identifies promising clauses for given clause selection,
while our recommender proposes a promising symbol precedence for the proof search.
The most recent version of ENIGMA, named ENIGMA Anonymous [Jakubův 2020],
uses the graph neural network (GNN) architecture proposed in [25],
and the relation of our work to this architecture is outlined in section 6 of our paper.
While both our recommender and ENIGMA use a GNN to process the input problems in a signature-agnostic manner,
the two GNNs are developed independently and their architectures are different.

- Despite the clear improvement over baseline, the percentage gained does not appear
to be too big. It is possible, that the potential effect of better orderings is not
too high, regardless of the optimizations. Can you give some insights for this
line of thought?

It is incredibly difficult to estimate, in absolute terms, how large an improvement
one could in our setting on TPTP achieve using some form of "perfect orderings".
However, experiments with random precedences [7] showed that the prover perceptibly responds 
to changes of this parameter and that work on a recommender is worthwhile.

[Jakubův 2020]: Jakubův, J., Chvalovský, K., Olšák, M., Piotrowski, B., Suda, M., Urban, J.: ENIGMA Anonymous: Symbol-Independent Inference Guiding Machine (System Description). In: LNCS. vol. 12167 LNAI, pp. 448–463. Springer (2020). https://doi.org/10.1007/978-3-030-51054-1_29

----------- Response to Review 2 -----------

- How does your tool behave in the presence of theories (assuming AVATAR is used)?

Both in the training and in the evaluation, AVATAR was disabled and we did not consider problems with theories.
Although, in principle, theory symbols could be subject to our recommender, Vampire treats them differently than the rest  
when constructing a precedence and we did not want to be sidetracked by these specificities.

- Now Vampire can also handle higher-order problems. Can your technique also improve Vampire's performance in this context?

Again, this seems in principle possible, as long as the ordering used by the HO vampire still relies on a precedence,
but we did not investigate the details yet.

- How long did it take to do all the training?

In total, the generation of the training data and the training of the model took approximately 8 days.
More precisely, it took approximately 2 and 4 days to generate the training data for the predicate and function precedence recommender, respectively,
and it took approximately 45 and 20 hours to train the predicate and function precedence recommender, respectively.

- What kind of new problems were solved in terms of the rating of the problems?

A total of 234 new problems were solved by the trained recommender when compared to the baseline heuristic.
Among these 234 problems, the maximum rating is 0.88,
the mean rating is approximately 0.36 and the median rating is 0.36.

- How independent are the training and test sets, given that the problems in TPTP are grouped by families that may include extremely similar problems?

In this experiment, we have not attempted to control nor analyze the similarity between the problems in the training and test sets.
Splitting the problem set by the domain or the source of the problems is, however, an interesting idea.

----------- Response to Review 3 -----------

- Why do you use the number or iterations of the main loop?

We mainly selected this metric because it is deterministic
and can be measured while other processes are making the computer busy.
Wall-clock time is more sensitive to the compute environment 
and more care needs to be taken to obtain meaningful results commeasureable across multiple runs.
However, the point raised by the reviewer is valid and suggests that using wall clock time as the decisive metric,
if implemented well, could increase the performance of the trained system.

- If I understand you correctly, the graph representation of a problem is fixed. So why don't you filter before generating training examples? CPU time must be cheap! ;-)

We chose to generate the training and validation examples (in the form of oriented pairs of precedences) independently from the graph representation of the problems.
This allowed us to adjust the hyperparameters of the two processes (example generation and graph representation) independently,
keeping the computation cost of the graph representation hyperparameter tuning relatively low.