### 2021-04-14 main.tex
Clean up the acknowledgements
Follow the recommendation from Hanka's email from 2021-04-14.

\documentclass[runningheads]{llncs}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}

\usepackage{etoolbox}

\newtoggle{LONG}
%\toggletrue{LONG}
\togglefalse{LONG}

\newtoggle{SHRINK}
%\toggletrue{SHRINK}
\togglefalse{SHRINK}

\usepackage[binary-units=true,detect-weight=true]{siunitx}

% How to write a paper:
% https://mj.ucw.cz/papers/jakpsat.pdf
% Jones 2016: https://www.microsoft.com/en-us/research/academic-program/write-great-research-paper/
% AWR: https://jazyky.fel.cvut.cz/vyuka/RPP/BE9M04AKP/

% PAAR paper: https://github.com/filipbartek/learning-precedences-from-elementary-symbol-features/releases/download/paar2020%2Fceur-2/Learning_Precedences_from_Simple_Symbol_Features.pdf

\input{preamble}
\input{glossary}
\input{front}

\usepackage{svg}

\usepackage[pdf]{graphviz}
\usepackage{tikz}

\usepackage{hyperref}
\hypersetup{
pdftitle={Neural Precedence Recommender},
pdfauthor={Filip Bártek, Martin Suda},
pdfkeywords={saturation-based theorem proving, simplification ordering, symbol precedence, machine learning, graph convolutional network}
}

\usepackage{cleveref}

\iftoggle{SHRINK}{
% Some mild shrinkage:

% for Section and subsection headings:
%% Save the class definition of \subparagraph
\let\llncssubparagraph\subparagraph
%% Provide a definition to \subparagraph to keep titlesec happy
\let\subparagraph\paragraph
%% Load titlesec
\usepackage{titlesec}
%% Revert \subparagraph to the llncs definition
\let\subparagraph\llncssubparagraph
\titlespacing*{\subsection}
  {0pt}{0.9\baselineskip}{0.8\baselineskip}
}

% For figures
% https://tex.stackexchange.com/questions/60477/remove-space-after-figure-and-before-text
\setlength{\textfloatsep}{0.7\baselineskip plus 0.2\baselineskip minus 0.2\baselineskip} % between a float at the top and text
\setlength{\floatsep}{0.3\baselineskip plus 0.2\baselineskip minus 0.2\baselineskip} % between two floats at the top
\setlength{\intextsep}{0.7\baselineskip plus 0.2\baselineskip minus 0.2\baselineskip} % between a float in text and text

\begin{document}

\maketitle

\begin{abstract}
\input{abstract}
\end{abstract}

\section{Introduction}

\input{introduction}

\section{Preliminaries}
\label{sec:preliminaries}
\input{preliminaries}

\section{Architecture}
\label{sec:architecture}
\input{architecture}

\section{Experimental evaluation}
\label{sec:evaluation}
\input{evaluation}

\section{Related work}
\label{sec:related}

Our previous text \cite{DBLP:conf/cade/Bartek020} marked the initial investigation of applying techniques of \acrlong{ml}
to generating good symbol precedences.
The neural recommender presented here uses a \gls{gnn} to model symbol costs,
while \cite{DBLP:conf/cade/Bartek020} used a linear combination of symbol features readily available in the \gls{atp} \Vampire{}.
The \gls{gnn}-based approach yields more performant precedences at the cost of longer training and preprocessing time.

In \cite{Olsak2019} and \cite{Rawson2020}, the authors propose similar \gls{gnn} architectures to solve tasks on \gls{fol} problems.
They use the \glspl{gnn} to solve classification tasks such as premise selection.
While our system is trained on a proxy classification task,
the main task it is evaluated on is the generation of useful precedences.

The problem of learning to rank objects represented by scores trainable by gradient descent was explored in \cite{Burges2005}.
Our work can be seen to apply the approach of \cite{Burges2005} to rank permutations represented by weighted sums of symbol costs.
\todo{Related problem: Ranking. See Mohri etc.}

\todo[inline,author=R1]{An obvious question is how is this work related to the work and methods
used for Enigma?

Although the paper related to the Enigma system by Jakubuv has been
cited, it appears to not be referred to in the text. It should be referred and
the reference list should be checked.
}

% Fast differentiable ranking (and probably other texts): They use the permutation scoring by weighted sum.

\section{Conclusion and future work}
\label{sec:conclusion}
\input{conclusion}

\section*{Acknowledgments}
% > Acknowledgements should generally be placed in an unnumbered subsection at the end of the paper.

% https://sgs.cvut.cz/index.php?action=faq
% This work was supported by the Grant Agency of the Czech Technical University in Prague, grant No. SGS...

This work was supported by
the Czech Science Foundation project no.~20-06390Y (JUNIOR grant),
the project RICAIP no. 857306 under the EU-H2020 programme,
and
the Grant Agency of the Czech Technical University in Prague, grant
no.~SGS20/215/OHK3/3T/37.

% > LaTeX users should avoid self-defined environments and use the bibliographic style MathPhySci for computer science proceedings.
% > It is not possible to have hyperlinks in references.
\bibliographystyle{splncs04}
\bibliography{main}

%\appendix
%\input{appendix}

\end{document}