### 2021-04-20 conclusion.tex
conclusion typos

% !TEX root = main.tex

We have described a system that extracts useful symbol precedences from the graph representations of \gls{cnf} problems.
Comparison against a conventional symbol precedence heuristic shows that using a \gls{gcn}
to consider the whole structure of the input problem is beneficial.
\todo{FB: Should we boast that the reduction of ranking to training 
of differentiable element scores is innovative? MS: Yes! Ideally (also) earlier.}

A manual analysis of the trained recommender could produce new insights into how the choice of the symbol precedence influences the proof search,
which could in turn help designing new efficient precedence generating schemes.
Indeed, a trained cost model summarizes the observed behaviours of the ATP on the random precedences
and is able to discover patterns in them (as we know implicitly from its accuracy) 
despite their seemingly chaotic behaviour as perceived by a human observer.
The challenge is to extract these patterns to a human-understandable form.
\todo[inline,author=R2]{[...] the analysis of the key patterns used by the trained recommender (only hinted at as a future work in the conclusion) would have been much more interesting in my opinion.}

In addition to the symbol precedence,
\gls{kbo} is determined by symbol \emph{weights}.
In this work, we kept the symbol weights fixed to the value 1.
Learning to recommend symbol weights in addition to the precedences
represents an interesting avenue for future research.

The same applies to the idea of learning to recommend 
both the predicate and the function precedences using a single \gls{gcn}.
The joint learning, although more complex to design, could 
additionally discover interdependencies
between the effects of the function precedence and predicate precedence on the proof search,
while the current setup implicitly assumes that the effects are independent.
Finally, a higher training data efficiency could be achieved by considering all pairs of measured executions on a problem
in one training batch.

\todo{MS: ja bych asi delsi zaver nepsal, ani kdyz bude misto...}

\iftoggle{LONG}{
A more detailed analysis of the training process might yield insights into how the performance generalizes
from the training set to the validation set
and from proxy metrics, such as the loss and the accuracy, to the final solver performance.
Ultimately, this might reveal modifications that make the training more efficient and the final recommender more performant.

The training examples take the form of pairs of precedences,
where one of the precedences is preferable over the other.
In this work, the amount by which the quality of the precedences differs is ignored.
Taking the exact or estimated numbers of saturation loop iterations into account
might help the recommender to prioritize the more significant gains.

Applying techniques of reinforcement learning to focus the training data on the cases that yield high performance gains
could further improve the recommender performance.

We restricted our experiments to problems whose graph representation does not exceed \num{100000} nodes.
Evaluating the recommender on larger problems would be an interesting test of generalization.
Training on larger problems may be facilitated by employing a lossy graphification.

We performed all experiments with a fixed \gls{atp} \Vampire{} and a fixed configuration
we consider to be a reasonable baseline.
Introducing other prover configurations (namely other \glspl{sot})
or other provers could make the resulting recommender more robust
and possibly help extract more general knowledge about the structure of good symbol precedences.

Finally, the approach outlined in \cref{sec:architecture}
could be adjusted to solve tasks on \gls{fol} formulas other than recommending symbol precedences,
such as premise selection of given clause selection.
On the other hand, we could try to solve other tasks that involve generating permutations of arbitrary length.
\todo{FB: TSP? Probably not because it is too combinatorial.}
}{}