### 2021-04-27 abstract.tex
Resolve overfull lines
by small changes.

% !TEX root = main.tex

The state-of-the-art superposition-based theorem provers for first order logic
rely on \glspl{sot} to constrain the applicability of inference rules,
which in turn shapes the ensuing search space.
The popular Knuth-Bendix simplification ordering is parameterized by 
\emph{symbol precedence}---a permutation of the predicate and function symbols
of the input problem's signature.
Thus, the choice of precedence has an indirect yet often substantial impact
on the amount of work required to complete a proof search successfully.

This paper describes and evaluates a symbol precedence recommender,
a \acrlong{ml} system that estimates the best possible precedence
based on observations of prover performance on a set of problems and random precedences.
Using the graph convolutional neural network technology,
the system does not presuppose the problems to be related or share a common signature. 
When coupled with the theorem prover \Vampire{} and evaluated on the \acrshort{tptp} problem library,
the recommender is found to outperform a state-of-the-art heuristic by more than \SI{4}{\percent}
on unseen problems.

\keywords{saturation-based theorem proving \and
simplification ordering \and
symbol precedence \and
\acrlong{ml} \and
\acrlong{gcn}}

% AWR: Guidelines (Unit 7, page 5):
% At most 300 words
% Content:
% Purpose of the study
% Research problems
% Basic design of the study
% Summary of interpretations and conclusions

% AWR: Parts of an abstract (checklist) (Unit 7, page 6):
% Motivation
% Problem statement
% Approach
% Results
% Conclusions

% Motivation and purpose of the study are excluded because the audience at CADE need not be reminded of the motivation for making the provers faster.