### 2021-04-27 main.tex
Move the content of section Related to a dedicated file

\documentclass[runningheads]{llncs}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}

\usepackage{etoolbox}

\newtoggle{LONG}
%\toggletrue{LONG}
\togglefalse{LONG}

\newtoggle{SHRINK}
%\toggletrue{SHRINK}
\togglefalse{SHRINK}

\usepackage[binary-units=true,detect-weight=true]{siunitx}

% How to write a paper:
% https://mj.ucw.cz/papers/jakpsat.pdf
% Jones 2016: https://www.microsoft.com/en-us/research/academic-program/write-great-research-paper/
% AWR: https://jazyky.fel.cvut.cz/vyuka/RPP/BE9M04AKP/

% PAAR paper: https://github.com/filipbartek/learning-precedences-from-elementary-symbol-features/releases/download/paar2020%2Fceur-2/Learning_Precedences_from_Simple_Symbol_Features.pdf

\input{preamble}
\input{glossary}
\input{front}

\usepackage{svg}

\usepackage[pdf]{graphviz}
\usepackage{tikz}

\usepackage{hyperref}
\hypersetup{
pdftitle={Neural Precedence Recommender},
pdfauthor={Filip Bártek, Martin Suda},
pdfkeywords={saturation-based theorem proving, simplification ordering, symbol precedence, machine learning, graph convolutional network}
}

\usepackage[capitalise]{cleveref}
\crefname{section}{Sect.}{Sects.}
\Crefname{section}{Section}{Sections}

\iftoggle{SHRINK}{
% Some mild shrinkage:

% for Section and subsection headings:
%% Save the class definition of \subparagraph
\let\llncssubparagraph\subparagraph
%% Provide a definition to \subparagraph to keep titlesec happy
\let\subparagraph\paragraph
%% Load titlesec
\usepackage{titlesec}
%% Revert \subparagraph to the llncs definition
\let\subparagraph\llncssubparagraph
\titlespacing*{\subsection}
  {0pt}{0.9\baselineskip}{0.8\baselineskip}
}

% For figures
% https://tex.stackexchange.com/questions/60477/remove-space-after-figure-and-before-text
\setlength{\textfloatsep}{0.7\baselineskip plus 0.2\baselineskip minus 0.2\baselineskip} % between a float at the top and text
\setlength{\floatsep}{0.3\baselineskip plus 0.2\baselineskip minus 0.2\baselineskip} % between two floats at the top
\setlength{\intextsep}{0.7\baselineskip plus 0.2\baselineskip minus 0.2\baselineskip} % between a float in text and text

\usepackage{times}

\begin{document}

\maketitle

\begin{abstract}
\input{abstract}
\end{abstract}

\section{Introduction}

\input{introduction}

\section{Preliminaries}
\label{sec:preliminaries}
\input{preliminaries}

\section{Architecture}
\label{sec:architecture}
\input{architecture}

\section{Experimental evaluation}
\label{sec:evaluation}
\input{evaluation}

\section{Related work}
\label{sec:related}
\input{related}

\section{Conclusion and future work}
\label{sec:conclusion}
\input{conclusion}

\section*{Acknowledgments}
% > Acknowledgements should generally be placed in an unnumbered subsection at the end of the paper.

% https://sgs.cvut.cz/index.php?action=faq
% This work was supported by the Grant Agency of the Czech Technical University in Prague, grant No. SGS...

This work was generously supported by
the Czech Science Foundation project no.~20-06390Y (JUNIOR grant),
the project RICAIP no.~857306 under the EU-H2020 programme,
and
the Grant Agency of the Czech Technical University in Prague, grant
no.~SGS20/215/OHK3/3T/37.

% > LaTeX users should avoid self-defined environments and use the bibliographic style MathPhySci for computer science proceedings.
% > It is not possible to have hyperlinks in references.
\bibliographystyle{splncs04}
\bibliography{main}

\end{document}